Von Neumann-Morgenstern and Savage Theorems for Causal Decision Making

### 2020-09-16 Artificial_Intelligence_Journal/Draft/AIJ_Draft.tex
cambios menores

\section{The Classical Decision Making Theorems: von Neumann-Morgenstern and Savage}
\label{classical_theorems}
A Decision Droblem under Uncertainty is the mathematical model of a situation in which an agent must choose one out of many available actions with uncertain consequences which depend on different, possibly unknown, factors \citep{bernardo2000bayesian,gilboa2009decision}. Such consequences are ordered in terms of the \textit{satisfaction} that they produce in the decision maker, and such ordering is represented by a \textit{preference relation} denoted by $\succeq$, where $a \succeq b$ is read as $a$ being preferred to $b$.

@@

\section{Application: Causal Games and Nash Equilibrium}
\label{causal_games}
In this section we know consider an application in the domain of game theory: we consider a \textit{strategic game} between $N$ rational players who are situated in a causal environment. We explore the implications of causal decision making in game theory given the importance of multi-agent systems for computational intelligence \cite{shoham2008multiagen,wooldridge2009introductiont}. With this contribution we expect to show that standard notions of game theory such as Nash Equilibrium can be extended to the case in which causal information is considered over associative information. Therefore, provide motivation to further extend classical results, which have been shown useful, to use causal information as a basis.

@@

In the method described in \cite{gonzalez2018playing}, the decision maker holds probabilistic beliefs about the parameters of the model, which are used in order to build a \textit{local} causal model that is used \textit{as if} it were the true one in order to obtain the \textit{best action} in each round. \cite{gonzalez2018playing}, assume that the graphical structure of the causal model is known and use a Dirichlet distribution in order to sample the Conditional Probability Tables required to fully specify a Causal Graphical Model; once such model is specified the authors use an expression derived from Pearl's do-calculus in order to find a best action; the best action is defined in terms of the maximum expected utility as required by both the von Neumann-Morgenstern and Savage Theorems. Their method fits within our Theorem \ref{causal_savage} because in this paper the distribution $P_C(g)$ is a \textit{distribution over causal models with the same structure}, and when a model is specified from $P_C$, it is then used as if it were the true one in order to find the best action for such causal model. In the above mentioned paper, the possible outcomes are only $1$ and $0$, so once a model $g$ is sampled, the product $u(c)P_g(c|do(a))$ simply becomes $P(C=1 | do(a))$ which is stated only as an intuitive solution in the mentioned paper. Our result Theorem \ref{causal_savage} formalizes their method, thus providing solid ground for learning algorithms which rely on causal information.

@@

%Queda la duda de si se necesita especificar un orden previo
\subsection{A random graph approach to causal structure learning}
\label{random_graphs}
For the sake of completeness, we now provide a relaxation of the restrictions found in \cite{lattimoreNIPS2016,sen2017identifying,gonzalez2018playing}: knowing the causal model, or some parts of it; in particular, we have mentione that in \cite{gonzalez2018playing} it is required that the graphical structure of the causal model is known, which is a strong assumption. Such assumption is now to be relaxed to the case in which the graphical structure of the causal model is unknown, but the variables of the model are known; also, we assume the decision maker has an ordering $\gg$ according to which if $A \gg B$, then $B$ can not cause $A$. Our extension will take on a purely Bayesian approach and make use both of Theorem \ref{causal_savage}, random graphs and the \textit{as if} intuition present in this paper.

@@

\section{Future Work}
In Section \ref{random_graphs} we considered the problem of a decision maker situated in a causal environment who does not fully knows the causal model which controls her environment; previous works have considered different levels of uncertainty over the causal model: knowing conditional distributions, knowing a part of the model, knowing the structure of the model or the variables involved. We leave as future work the case in which the variables of the model are not known and we consider that this scenario could be attacked using tools from Bayesian Nonparametrics \citep{hjort2010bayesian,phadia2013prior} in the form of a prior over causal models, which generates causal graphs of distinct sizes and complexity.

>>>>

\section{The Classical Decision Making Theorems: von Neumann-Morgenstern and Savage}
\label{classical_theorems}
A Decision Problem under Uncertainty is the mathematical model of a situation in which an agent must choose one out of many available actions with uncertain consequences which depend on different, possibly unknown, factors \citep{bernardo2000bayesian,gilboa2009decision}. Such consequences are ordered in terms of the \textit{satisfaction} that they produce in the decision maker, and such ordering is represented by a \textit{preference relation} denoted by $\succeq$, where $a \succeq b$ is read as $a$ being preferred to $b$.

@@

\section{Application: Causal Games and Nash Equilibrium}
\label{causal_games}
In this section we know consider an application in the domain of game theory: we consider a \textit{strategic game} between $N$ rational players who are situated in a causal environment. We explore the implications of causal decision making in game theory given the importance of multi-agent systems for computational intelligence \citep{shoham2008multiagent,wooldridge2009introduction}. With this contribution we expect to show that standard notions of game theory such as Nash Equilibrium can be extended to the case in which causal information is considered over associative information. Therefore, provide motivation to further extend classical results, which have been shown useful, to use causal information as a basis.

@@

In the method described in \cite{gonzalez2018playing}, the decision maker holds probabilistic beliefs about the parameters of the model, which are used in order to build a \textit{local} causal model that is used \textit{as if} it were the true one in order to obtain the \textit{best action} in each round. \cite{gonzalez2018playing}, assume that the graphical structure of the causal model is known and use a Dirichlet distribution in order to sample the Conditional Probability Tables required to fully specify a Causal Graphical Model; once such model is specified the authors use an expression derived from Pearl's do-calculus in order to find a best action; the best action is defined in terms of the maximum expected utility as required by both the von Neumann-Morgenstern and Savage Theorems. Their method fits within our Theorem \ref{causal_savage} because in this paper the distribution $P_C(g)$ is a \textit{distribution over causal models with the same structure}, and when a model is specified from $P_C$, it is then used as if it were the true one in order to find the best action for such causal model. In the above mentioned paper, the possible outcomes are only $1$ and $0$, so once a model $g$ is sampled, the product $u(c)P_g(c|do(a))$ simply becomes $P(C=1 | do(a))$ which is stated only as an intuitive solution in the mentioned paper. Our result, Theorem \ref{causal_savage}, formalizes their method, thus providing solid ground for learning algorithms which rely on causal information.

@@

%Queda la duda de si se necesita especificar un orden previo
\subsection{A random graph approach to causal structure learning}
\label{random_graphs}
For the sake of completeness, we now provide a relaxation of the restrictions found in \cite{lattimoreNIPS2016,sen2017identifying,gonzalez2018playing}: knowing the causal model, or some parts of it; in particular, as stated before in \cite{gonzalez2018playing}, it is required that the graphical structure of the causal model is known, which is a strong assumption. Such assumption is now to be relaxed to the case in which the graphical structure of the causal model is unknown, but the variables of the model are known; also, we assume the decision maker has an ordering $\gg$ according to which if $A \gg B$, then $B$ can not cause $A$. Our extension will take on a purely Bayesian approach and make use both, Theorem \ref{causal_savage} and random graphs as well as the \textit{as if} intuition present in this paper.

@@

\section{Future Work}
We have left out the implementation of the algorithm stated in Section \ref{random_graphs} we considered the problem of a decision maker situated in a causal environment who does not fully knows the causal model which controls her environment; previous works have considered different levels of uncertainty over the causal model: knowing conditional distributions, knowing a part of the model, knowing the structure of the model or the variables involved. We leave as future work the case in which the variables of the model are not known and we consider that this scenario could be attacked using tools from Bayesian Nonparametrics \citep{hjort2010bayesian,phadia2013prior} in the form of a prior over causal models. If only a part of the set of variables is known, then a Bayesian Nonparametric prior, which generates causal graphs of distinct sizes and complexity, could capture the patterns of causal interaction between known and possible but unknown variables.
